maxent model avec xvalidation (year-based test training)

### 2019-07-15 src/maxent_alldata_avec_xvalidation.Rmd
fecky feck violin plots

```{r pre-install & load packages, include=FALSE}
options (java.parameters = "-Xmx2048m")
packages <- c("rJava", "dismo", "stringr", "raster", "ecospat")
source("../src/ipak.R")
ipak(packages)
```

@@

```{r}
allvar <- c("occurrence", "year", "month", "temp_surface", "temp_depth", "salinity_surface", "salinity_depth", "o2_depth", "chl_surface", "mlp_surface",  "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
prback <- subset(prback, select = c(allvar))
colnames(prback)
```

@@

```{r remove na vals}
prback <- na.omit(prback)

prback <- subset(prback, month == 8)

presyrs <-  subset(prback, occurrence == 1) #this is to get years with complete presence data 
presyrs <- unique(presyrs$year) #make a vector of unique years
prback <- prback[prback$year %in% presyrs, ] #keep only the data for which there is precence years
```

@@

```{r train and test inc AIC for reg val}
aicetc <- read.csv("../data/env/AICcCalculations.csv", header = TRUE) #pre-created csv file
#betavallist <- cbind(aicetc$regularization)
betavallist <- seq(0.25, 0.75, by = 0.25)

timeslice <- "all_data"
train <- prback
trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
nfolds <- cbind(c(1:2))

 for (h in 1:length(betavallist)){
   betamval <-  betavallist[h]
    #betamval <-  0.5
    mod <- maxent(x = train, p = trainpa, plots = FALSE, path = paste0("../output/maxent/", timeslice, "/models"), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("replicates=", nrow(nfolds)), "replicatetype=crossvalidate", paste0("betamultiplier=", betamval)))

@@

    aicperbetalst <- list.files(paste0("../output/maxent/", timeslice, "/models/"), pattern = paste0(".*AICforMaxentReg.csv"), full.names = TRUE)
    for (beta in 1:length(aicperbetalst)){
      aic <- read.csv(aicperbetalst[beta])
      bval <- str_sub(aicperbetalst[beta], 35, -20) #the year being used for testing
      aicmean <- colwise(mean)(aic)
      write.csv(aicmean, paste0("../output/maxent/", timeslice, "/models/aic_mean_", bval, ".csv"), row.names = FALSE)
    }

@@

#model selection
next step... select the models you want for each year and run those models again (primarily based on lowest AICc but check test AUC and other things too)

KEEP A NOTE OF THESE IN ONENOTE

y1998 = 2.75
y1999 = 2.25
y2000 = 2
y2001 = 1.75
y2002 = 2
y2003 = 2
y2004 = 2
y2005 = 2.25
y2006 = 1.75
y2007 = 1.75
y2008 = 2.25
y2009 = 2.25
y2010 = 2.5
y2011 =  2.5

```{r final models plus predict}
modelmonth <- cbind(paste0(mthna, unique(prback$year)))
aicetc <- read.csv("../data/env/AICcFinalmod.csv", header = TRUE) #pre-created csv file
unlayer <- read.csv(paste0("../data/env/asc_layerlist", mthna, ".csv"), header = TRUE) #if code doesnt work put back to unlayer_yr

#for (year in 1:nrow(modelmonth)){
 # timeslice <- modelmonth[year] #testing timeslice
  #yr <- str_sub(timeslice, -4) #the year being used for testing
  train <- subset(prback, year != yr) #training data is all years but testing
  trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
  test <-  subset(prback, year == yr) #creates the testing dataset
  testp <- subset(test, occurrence == "1") #splits the testing dataset into presences
  testp <- subset(testp, select = -c(year, occurrence)) #don't need year and occurrence for modelling
  testb <- subset(test, occurrence == "0") #splits the testing dataset into background
  testb <- subset(testb, select = -c(year, occurrence)) #don't need year and occurrence for modelling
  train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling

  if (yr == "2003" | yr == "2004"){
    betaval <- 1
  } else if (yr == "2005"){
    betaval <- 1.5
  } else if (yr == "2006"){
    betaval <- 2.25
  }

    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betaval)))

@@

      aicetc$train_AUC <- trainauc
      aicetc$no_trainng_points <- notrainingpoints
      aicetc$test_coefficients <- nocoeff
      aicetc$test_sum.of.logs <- sumsamplepred
      aicetc$test_AUC <- evauc
      aicetc$test_evcor <- evcor
      aicetc$test_evpcor <- evpcor
      aicetc$test_nopresence <- evnp
      aicetc$test_nobackground <- evna
      aicetc$test_maxsumsenspec <- evmaxtprtnr
      aicetc$test_maxkap <- evmaxkap
      aicetc$test_no_omission <- evno_omission
      aicetc$test_prevalence <- evprevalence
      aicetc$test_equal_sens_spec <- evequal_sens_spec
      aicetc$test_sensitivity <- evsensitivity
      aicetc$test_bic <- bic #not sure this is correct...
      aicetc$beta <- betaval

    aicetc$AIC <- ((2*aicetc$coefficients)-(2*aicetc$sum.of.logs))
    aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$n/(aicetc$n - aicetc$coefficients-1))))

    #calculate minimal predicted area using ecospat
    aicetc$mpa_0.9_cum <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
    aicetc$mpa_0.9_clog <- ecospat.mpa(samplepred$Cloglog.prediction, perc = 0.9) #90%

    write.csv(aicetc, paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)

    unlayer_yr <- subset(unlayer, ascyears == yr)
    unlayer_yr <- unlayer_yr[-c(1), ]
    unlayer_yr[] <- lapply(unlayer_yr, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
    stackdeplst <- list()
    stackdep <- subset(unlayer_yr, select = c(stackname, ascdepths))
    stackdeplst[[year]] <- stackdep
    #write.csv(stackdep, paste0("../output/maxent/predictions/stackdepthlist_", yr, ".csv"), row.names = FALSE)
    aea <- raster("../output/env/aea.tif")

   for (td in 1:nrow(unlayer_yr)){
      temp_depth <- raster(unlayer_yr$temp_depth[td])
      crs(temp_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(temp_depth) <- "temp_depth"
      temp_depth <- projectRaster(temp_depth, aea)
      chl_surface <- raster(unlayer_yr$chl_surface[td])
      crs(chl_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(chl_surface) <- "chl_surface"
      chl_surface <- projectRaster(chl_surface, aea)
      o2_depth <- raster(unlayer_yr$o2_depth[td])
      crs(o2_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(o2_depth) <- "o2_depth"
      o2_depth <- projectRaster(o2_depth, aea)
      salinity_depth <- raster(unlayer_yr$salinity_depth[td])
      crs(salinity_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(salinity_depth) <- "salinity_depth"
      salinity_depth <- projectRaster(salinity_depth, aea)
      nao_sample <- raster(unlayer_yr$nao_sample[td])
      crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(nao_sample) <- "nao_sample"
      nao_sample <- projectRaster(nao_sample, aea)
      nao_winter <- raster(unlayer_yr$nao_winter[td])
      crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
      names(nao_winter) <- "nao_winter"
      nao_winter <- projectRaster(nao_winter, aea)
      stackn <- unlayer_yr$stackname[td]
      stack <- stack(temp_depth, salinity_depth, o2_depth, chl_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_winter)
      #stackSave(test,unlayer_yr$stackname[td])
      pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", "cloglog_", stackn, ".asc"), overwrite=TRUE, args='outputformat=cloglog') 
      predcum <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", "cumu_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=Cumulative')) 
      prplot <- plot(pred, png(paste0("../output/maxent/predictions/", mthna, "/", "cloglog_", stackn, ".png")))
      dev.off() # stops automatic saving of the plot to a png

@@

stdep <- do.call(rbind, stackdeplst)
write.csv(stdep, paste0("../output/maxent/predictions/", mthna, "/stack_depth.csv"), row.names = FALSE)
```

@@

```{r avg maxentReg}

for (i in 1:length(modelmonth)){
  timeslice <- modelmonth[i]
  yr <- str_sub(timeslice, -4) #the year being used for testing
  if(yr == 1998){
    aic1998 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr1998 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 1999){
    aic1999 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr1999 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2000){
    aic2000 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2000 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2001){
    aic2001 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2001 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2002){
    aic2002 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2002 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2003){
    aic2003 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2003 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2004){
    aic2004 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2004 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2005){
    aic2005 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2005 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2006){
    aic2006 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2006 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2007){
    aic2007 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2007 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2008){
    aic2008 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2008 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2009){
    aic2009 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2009 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2010){
    aic2010 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2010 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2011){
    aic2011 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2011 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2012){
    aic2012 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2012 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2013){
    aic2013 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2013 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2014){
    aic2014 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2014 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2015){
    aic2015 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2015 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
  }
}

aic <- rbind(aic2003, aic2004, aic2005, aic2006)
maxr <- rbind(maxr2003, maxr2004, maxr2005, maxr2006)

aicm <- colMeans(aic)
aicm$mod <- "avg"
avgAIC <- rbind(aic, aicm)

maxr$Species <- 1
maxrm <- colMeans(maxr)
maxrm$Species <- "avg"
maxrm <- rbind(maxr, maxrm)

write.csv(avgAIC, paste0("../output/maxent/models/", mthna, "avgAIC.csv"), row.names = FALSE)
write.csv(maxrm, paste0("../output/maxent/models/", mthna, "avgmaxentresults.csv"), row.names = FALSE)
```

@@

```{r avg asc layers cloglog}
asc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "cloglog*.asc", full.names = TRUE)
depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
names(depthlayers)[names(depthlayers) == "cbind.unique.stdep.ascdepths.."] <- "depth"
depthlayers$depth <- as.character(depthlayers$depth)

@@

for (d in 1:nrow(depthlayers)){
  dep <- depthlayers$depth[d]
  for (a in 1:length(asc_list)){
    asc <- asc_list[a]
    if (grepl(dep, asc)){
      depthlayers[d , paste0("a", colno)] <- asc
        colno <- colno+1
    }
  }
}
write.csv(depthlayers, (paste0("../output/maxent/predictions/", mthna, "/cloglog_asclayerbydepth.csv")), row.names = TRUE)

for (x in 1:nrow(depthlayers)){
  asclay <- depthlayers[x, ]
  asclay <- asclay[!is.na(asclay)]
  asclay <- as.data.frame(rbind(asclay))
  asclay$V1 <- as.character(asclay$V1) #v1 corresponds to depth
  asclay$V2 <- as.character(asclay$V2) #v2 to end corresponds to year
  asclay$V3 <- as.character(asclay$V3)
  asclay$V4 <- as.character(asclay$V4)
  asclay$V5 <- as.character(asclay$V5)
  asclay$V6 <- as.character(asclay$V6)
  asclay$V7 <- as.character(asclay$V7)
  asclay$V8 <- as.character(asclay$V8)
  asclay$V9 <- as.character(asclay$V9)
  asclay$V10 <- as.character(asclay$V10)
  asclay$V11 <- as.character(asclay$V11)
  asclay$V12 <- as.character(asclay$V12)
  asclay$V13 <- as.character(asclay$V13)
  asclay$V14 <- as.character(asclay$V14)
  asclay$V15 <- as.character(asclay$V15)

  asclaydep <- asclay$V1

  asc1998 <- raster(asclay$V2)
  asc1999 <- raster(asclay$V3)
  asc2000 <- raster(asclay$V4)
  asc2001 <- raster(asclay$V5)
  asc2002 <- raster(asclay$V6)
  asc2003 <- raster(asclay$V7)
  asc2004 <- raster(asclay$V8)
  asc2005 <- raster(asclay$V9)
  asc2006 <- raster(asclay$V10)
  asc2007 <- raster(asclay$V11)
  asc2008 <- raster(asclay$V12)
  asc2009 <- raster(asclay$V13)
  asc2010 <- raster(asclay$V14)
  asc2011 <- raster(asclay$V15)

  stkasc <- stack(asc1998, asc1999, asc2000, asc2001, asc2002, asc2003, asc2004, asc2005, asc2006, asc2007, asc2008, asc2009, asc2010, asc2011)
  avg <- calc(stkasc, fun = mean)

  writeRaster(avg, (paste0("../output/maxent/predictions/", mthna, "/avg_", asclaydep, ".asc")), overwrite=TRUE, args='outputformat=cloglog')
  avgplot <- plot(avg, png(paste0("../output/maxent/predictions/", mthna, "/avg_", asclaydep, ".png"))) 
  dev.off()
}

ascdepth_list <- list.files(paste0("../output/maxent/predictions/", mthna), glob2rx(pattern = "avg_*.asc"), full.names = TRUE)

clog_asc120  <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_120.asc"))
clog_asc41.18 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_41.18.asc"))
clog_asc1.55586 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1.55586.asc"))
clog_asc1045.85 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1045.85.asc"))
clog_asc108.03 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_108.03.asc"))
clog_asc11.7737 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_11.7737.asc"))
clog_asc1151.99 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1151.99.asc"))
clog_asc1265.86 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1265.86.asc"))
clog_asc13.991 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_13.991.asc"))
clog_asc133.076 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_133.076.asc"))
clog_asc1387.38 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_1387.38.asc"))
clog_asc147.406 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_147.406.asc"))
clog_asc16.5253 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_16.5253.asc"))
clog_asc163.165 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_163.165.asc"))
clog_asc180.55 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_180.55.asc"))
clog_asc19.4298 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_19.4298.asc"))
clog_asc199.79 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_199.79.asc"))
clog_asc2.66768 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_2.66768.asc"))
clog_asc22.7576 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_22.7576.asc"))
clog_asc221.141 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_221.141.asc"))
clog_asc244.891 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_244.891.asc"))
clog_asc26.5583 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_26.5583.asc"))
clog_asc271.356 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_271.356.asc"))
clog_asc3.85628 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_3.85628.asc"))
clog_asc30.8746 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_30.8746.asc"))
clog_asc300.888 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_300.888.asc"))
clog_asc333.863 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_333.863.asc"))
clog_asc35.7402 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_35.7402.asc"))
clog_asc370.689 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_370.689.asc"))
clog_asc411.794 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_411.794.asc"))
clog_asc457.626 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_457.626.asc"))
clog_asc47.2119 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_47.2119.asc"))
clog_asc5.14036 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_5.14036.asc"))
clog_asc508.64 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_508.64.asc"))
clog_asc53.8506 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_53.8506.asc"))
clog_asc565.292 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_565.292.asc"))
clog_asc6.54303 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_6.54303.asc"))
clog_asc61.1128 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_61.1128.asc"))
clog_asc628.026 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_628.026.asc"))
clog_asc69.0217 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_69.0217.asc"))
clog_asc697.259 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_697.259.asc"))
clog_asc77.6112 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_77.6112.asc"))
clog_asc773.368 <- raster(paste0("../output/maxent/predictions/", mthna, "/avg_773.368.asc"))

>>>>

```{r pre-install & load packages, include=FALSE}
gc() 
options (java.parameters = "-Xmx6g")
packages <- c("rJava", "dismo", "stringr", "raster", "ecospat", "plyr", "data.table", "RColorBrewer")
source("../src/ipak.R")
ipak(packages)
```

@@

```{r}
allvar <- c("occurrence", "year", "month", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
prback <- subset(prback, select = c(allvar))
colnames(prback)
```

@@

```{r remove na vals}
prback <- na.omit(prback)
#prback <- subset(prback, month == 8)
presyrs <-  subset(prback, occurrence == 1) #this is to get years with complete presence data 
presyrs <- unique(presyrs$year) #make a vector of unique years
prback <- prback[prback$year %in% presyrs, ] #keep only the data for which there is precence years
```

@@

```{r train and test inc AIC for reg val}
aicetc <- read.csv("../data/env/AICcCalculations.csv", header = TRUE) #pre-created csv file
betavallist <- seq(1.75, 4, by = 0.25)

timeslice <- "all_data"
train <- prback
trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
train  <- subset(train, select = -c(year, occurrence, month)) #don't need year and occurrence for modelling
nfolds <- cbind(c(1:5))

 for (h in 1:length(betavallist)){
   betamval <-  betavallist[h]
    #betamval <-  0.5
    mod <- maxent(x = train, p = trainpa, plots = FALSE, path = paste0("../output/maxent/", timeslice, "/models"), args=c("maximumiterations=5000", "responsecurves=FALSE", "jackknife=FALSE", "plots=FALSE", "threads=8", "addsamplestobackground=TRUE", paste0("replicates=", nrow(nfolds)), "replicatetype=crossvalidate", paste0("betamultiplier=", betamval)))

@@

    aicperbetalst <- list.files(paste0("../output/maxent/", timeslice, "/models/"), pattern = paste0(".*AICforMaxentReg.csv"), full.names = TRUE)
    for (beta in 1:length(aicperbetalst)){
      aic <- read.csv(aicperbetalst[beta])
      bval <- str_sub(aicperbetalst[beta], 35, -20) #the beta value
      aicmean <- colwise(mean)(aic)
      write.csv(aicmean, paste0("../output/maxent/", timeslice, "/models/aic_mean_", bval, ".csv"), row.names = FALSE)
    }

@@

#model selection

lowest aicc is 0.25

```{r final models plus predict}

aicetc <- read.csv("../data/env/AICcCalculations.csv", header = TRUE) #pre-created csv file
betaval <-  0.25

timeslice <- "all_data"
train <- prback
trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
train  <- subset(train, select = -c(year, occurrence, month)) #don't need year and occurrence for modelling
nfolds <- cbind(c(1:5))

mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/all_data/models"), args=c("maximumiterations=5000", "threads=8", "addsamplestobackground=TRUE", paste0("replicates=", nrow(nfolds)), "replicatetype=crossvalidate", paste0("betamultiplier=", betaval)))

for (i in 1:nrow(aicetc)){   
for(f in 1:nrow(nfolds)){
    modn <- nfolds[f]-1
      lamf <- lambdas_list[f]
      sampf <- samppred_list[f]
      coefile <- read.csv(paste0(lamf), header = FALSE, stringsAsFactors=FALSE)
      coefile <- head(coefile, -4)
      coefile2 <- subset(coefile, V2 != 0)
      nocoeff <- nrow(coefile2)
      bick <- sum(as.numeric(coefile2[, 2])) #dodgy 
      bic <- (bick*log(sum(trainpa == 1)) - 2*11) #dodgy

@@

    #continuous boyce index
    predPres <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
    predBg <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_backgroundPredictions.csv"), header = TRUE)
    predPres <- predPres$Cloglog.prediction
    predBg <- predBg$Cloglog
    cbi <- contBoyce(pres = predPres, bg = predBg)

      aicetc$train_AUC <- trainauc #training only AUC - not tested
      aicetc$no_trainng_points <- notrainingpoints #numberof points used for training
      aicetc$coefficients <- nocoeff #this is to calculate the AIC
      aicetc$sum.of.logs <- sumsamplepred #this is to calculate the AIC
      aicetc$test_AUC <- evauc #testing AUC
      aicetc$test_evcor <- evcor #testing correlation coefficient
      aicetc$test_evpcor <- evpcor # testing p-value for correlation coefficient
      aicetc$test_nopresence <- evnp 
      aicetc$test_nobackground <- evna
      aicetc$test_maxsumsenspec <- evmaxtprtnr
      aicetc$test_maxkap <- evmaxkap
      aicetc$test_no_omission <- evno_omission
      aicetc$test_prevalence <- evprevalence
      aicetc$test_equal_sens_spec <- evequal_sens_spec
      aicetc$test_sensitivity <- evsensitivity
      aicetc$tss <- max(ev@TPR + ev@TNR) - 1
      aicetc$test_bic <- bic #not sure this is correct...
      aicetc$beta <- betaval
      aicetc$cbi <- cbi

    aicetc$AIC <- ((2*aicetc$coefficients)-(2*aicetc$sum.of.logs))
    aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$no_trainng_points/(aicetc$no_trainng_points - aicetc$coefficients-1))))

    #calculate minimal predicted area using ecospat
    aicetc$mpa_0.9_cum <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
    aicetc$mpa_0.9_clog <- ecospat.mpa(samplepred$Cloglog.prediction, perc = 0.9) #90%
    aicetc$year <- yr

    write.csv(aicetc, paste0("../output/maxent/", timeslice, "/models/", "_", betaval, "_AICforMaxentReg.csv"), row.names = FALSE)

    saveRDS(mod, file = "../output/maxent/all_data/maxmod.rds", ascii = FALSE, version = NULL, compress = TRUE, refhook = NULL)

```{r final models predict}

filenames <- list.files("../data/all_data_layers/", pattern="*.csv", full.names=TRUE)
unlayer_yrX <- rbindlist(lapply(filenames,fread))
unlayer_yrX <- unlayer_yrX[complete.cases(unlayer_yrX), ]
aea <- raster("../output/env/aea.tif") 
unlayer_yr <- subset(unlayer_yrX, ascmonths == "7")

   for (td in 1:nrow(unlayer_yr)){
temp_depth <- raster(unlayer_yr$temp_depth[td])
      crs(temp_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(temp_depth) <- "temp_depth"
      temp_depth <- projectRaster(temp_depth, aea)
      chl_surface <- raster(unlayer_yr$chl_surface[td])
      crs(chl_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(chl_surface) <- "chl_surface"
      chl_surface <- projectRaster(chl_surface, aea)
      o2_depth <- raster(unlayer_yr$o2_depth[td])
      crs(o2_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(o2_depth) <- "o2_depth"
      o2_depth <- projectRaster(o2_depth, aea)
      salinity_depth <- raster(unlayer_yr$salinity_depth[td])
      crs(salinity_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(salinity_depth) <- "salinity_depth"
      salinity_depth <- projectRaster(salinity_depth, aea)
      nao_sample <- raster(unlayer_yr$nao_sample[td])
      crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(nao_sample) <- "nao_sample"
      nao_sample <- projectRaster(nao_sample, aea)
      nao_prev <- raster(unlayer_yr$nao_prev[td])
      crs(nao_prev) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(nao_prev) <- "nao_prev"
      nao_prev <- projectRaster(nao_prev, aea)
      nao_winter <- raster(unlayer_yr$nao_winter[td])
      crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
      names(nao_winter) <- "nao_winter"
      nao_winter <- projectRaster(nao_winter, aea)
      amo_sample <- raster(unlayer_yr$amo_sample[td])
      crs(amo_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(amo_sample) <- "amo_sample"
      amo_sample <- projectRaster(amo_sample, aea)
      amo_winter <- raster(unlayer_yr$amo_winter[td])
      crs(amo_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
      names(amo_winter) <- "amo_winter"
      amo_winter <- projectRaster(amo_winter, aea)
      stackn <- unlayer_yr$stackname[td]
      stack <- stack(temp_depth, salinity_depth, o2_depth, chl_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_winter)

      #stackSave(test,unlayer_yr$stackname[td])
      pred <- predict(mod, stack, file = paste0("../output/maxent/all_data/predictions/", "cloglog_", stackn, ".asc"), overwrite=TRUE, args='outputformat=cloglog') 
      #prplot <- plot(pred, png(paste0("../output/maxent/all_data/predictions/", "cloglog_", stackn, ".png")))
      #dev.off() # stops automatic saving of the plot to a png

@@

@@

@@

```{r avg asc layers cloglog}
stdep <- read.csv("../output/maxent/all_data/stackdepthlist.csv", header = TRUE)
asc_list <- list.files(paste0("../output/maxent/all_data/predictions/"), pattern = "cloglog.*.asc", full.names = TRUE)
depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
names(depthlayers)[names(depthlayers) == "cbind.unique.stdep.ascdepths.."] <- "depth"
depthlayers$depth <- as.character(depthlayers$depth)

@@

for (d in 1:nrow(depthlayers)){
  dep <- depthlayers$depth[d]
  for (a in 1:length(asc_list)){
    asc <- asc_list[a]
    if (grepl(dep, asc)){
      depthlayers[d , paste0("a", colno)] <- asc
      colno <- colno+1
    }
  }
}
write.csv(depthlayers, ("../output/maxent/all_data/predictions/cloglog_asclayerbydepth.csv"), row.names = FALSE)

#depthlayers <- read.csv("../output/maxent/all_data/predictions/cloglog_asclayerbydepth.csv", header = FALSE)

for (x in 1:nrow(depthlayers)){
  dstack <- stack()
  asclay <- depthlayers[x, ]
  asclay <- asclay[!is.na(asclay)]
  dep <- as.character(asclay[1])
  asclay <- as.data.frame(rbind(asclay))
  asclay[1] <- NULL
  for (ly in 1:length(asclay)){
    lay <- as.character(asclay[ ,40])
    layr <- raster(lay)
    dstack <- stack(dstack, layr)
    if(ly == length(asclay)){
      stavg <- calc(dstack, fun = mean, na.rm = TRUE)
      writeRaster(stavg, (paste0("../output/maxent/all_data/predictions/avg_", dep, ".asc")), overwrite=TRUE)
      avgplot <- plot(stavg, png(paste0("../output/maxent/all_data/predictions/avg_", dep, ".png")))
      while (!is.null(dev.list()))  dev.off()
    }
  }
}
